### 2014-09-19 paper/include/accessrecency.tex
cite added
explained techniques

smart
refrint
P. Emma et al.

>>>>

smart\cite{smartrefresh}

By employing a time-out counter for each memory row of a DRAM module, all the unnecessary periodic refresh operations can be eliminated. The basic concept behind our scheme is that a DRAM row that was recently read or written to by the processor (or other devices that share the same DRAM) does not need to be refreshed again by the periodic refresh operation, thereby eliminating excessive refreshes and the energy dissipated.

refrint\cite{refrint}

P. Emma et al.

### 2014-09-19 paper/include/retention.tex
added text and excluded pdf from git
cite added
explained techniques

\section{Retention}
\label{sec:ret}
\cite{ACM:dtail}

>>>>

\section{Retention}
\label{sec:ret}
Papers that use this technique:

DTail\cite{dtail}

VRA

RAIDR\cite{raidr}, from 2006...
Software approach that can exploit off-the-shelf DRAMs to reduce refresh power to vanishingly
small levels approaching non-volatile memory. The key idea is to favor longer-retention pages over shorter-retention pages when allocating DRAM pages. This allows selecting a single refresh period that depends on the shortest-retention page among populated pages,
instead of the shortest-reten tion page overall.

RIO\cite{rioparis}

SECRET\cite{secret}
Previous works adopt conventional ECC (Error Correcting Code) to correct retention errors. These approaches introduce significant area and energy overheads. In this paper, we propose a novel error correction framework for retention errors in DRAMs, called SECRET (Selective Error Correction for Refresh Energy reducTion).

Energy Optimization in 3D MPSoCs with Wide-I/O DRAM Using Temperature Variation Aware Bank-wise Refresh

We improve DRAM refresh power by considering the lateral and vertical temperature
variations in the 3D structure and adapting the per-DRAM-bank refresh period accordingly.

### 2014-09-19 paper/include/tolerance.tex
added text and excluded pdf from git
explained techniques

\section{Retention}
\label{sec:ret}

@@

DTail
VRA
RAIDR
RIO

>>>>

\section{Tolerance} 
\label{sec:tol}

@@

Flikker\cite{flikker}

Sparkk\cite{sparkk}
Multiple authors have suggested techniques where programmers can specify which data is critical and can not tolerate any bit errors and which data can be stored approximately. However, in these approaches all bits in the approximate area are treated as equally important. We show that this produces suboptimal results and higher energy savings or better quality can be achieved, if a more fine-grained approach is used. Our proposal is able to save more refresh power and enables a more effective storage of non-critical data by utilizing a non-uniform refresh of multiple DRAM chips and a permutation of the bits to the DRAM chips.
